---
title: "C67 a1"
output: pdf_document
date: "2022-10-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
library(tidyverse)
library(gridExtra)
```


# Question 1A


```{r}
set.seed(1006740421)
a=rnorm(100,mean=100,sd=20)
df1 = data.frame(a)
b=rnorm(1000,mean=100,sd=20)
df2 = data.frame(b)
c=rnorm(10000,mean=100,sd=20)
df3 = data.frame(c)
d=rnorm(100000,mean=100,sd=20)
df4 = data.frame(d)

```

```{r}

p1<-ggplot(df1,aes(a))+geom_histogram(bins=10,aes(y=..density..))+geom_density(col="blue")

p2<-ggplot(df2,aes(b))+geom_histogram(bins=10,aes(y=..density..))+geom_density(col="blue")

p3<-ggplot(df3,aes(c))+geom_histogram(bins=10,aes(y=..density..))+geom_density(col="blue")

p4<-ggplot(df4,aes(d))+geom_histogram(bins=10,aes(y=..density..))+geom_density(col="blue")

grid.arrange(p1,p2,p3,p4,nrow=2)

```

As the size increasing, the data will be closer to a normal distribution.

# Question 1B

```{r}
summary(a)
sd(a)
quantile(a, c(0.025, 0.25, 0.5, 0.75, 0.975))
```

```{r}
summary(b)
sd(b)
quantile(b, c(0.025, 0.25, 0.5, 0.75, 0.975))
```

```{r}
summary(c)
sd(c)
quantile(c, c(0.025, 0.25, 0.5, 0.75, 0.975))
```

```{r}
summary(d)
sd(d)
quantile(d, c(0.025, 0.25, 0.5, 0.75, 0.975))
```

```{r}
qnorm(c(.025, .25, .5, .75, .975),mean=100,sd=20)
```

For mean, the theoretical value is 100. 

Sample size = 100, mean = 101.65. 

Sample size = 1000, mean = 100.22. 

Sample size = 10000, mean = 100.24.

Sample size = 10000, mean = 100.10.


For sd, the theoretical value is 20.

Sample size = 100, sd = 20.88165. 

Sample size = 1000, sd = 20.2446. 

Sample size = 10000, sd = 20.02683. 

Sample size = 100000, sd = 19.99937.


For percentile (2.5, 25, 50, 75, 97.5), the theoretical value is (60.80072  86.51020 100.00000 113.48980 139.19928).

Sample size = 100, percentile (2.5, 25, 50, 75, 97.5) = (63.05151  87.42281 103.07220 114.84038 143.84550).

Sample size = 1000, percentile (2.5, 25, 50, 75, 97.5) = (60.22967  86.19834  99.87416 113.77874 136.93728).

Sample size = 10000, percentile (2.5, 25, 50, 75, 97.5) = (60.84855  86.77171 100.33805 113.80948 139.55371).

Sample size = 100000, percentile (2.5, 25, 50, 75, 97.5) = (60.9041  86.5363 100.1571 113.5054 139.4848).


We can make a conclusion that as the sample size increasing, the mean, sd, percentile (2.5, 25, 50, 75, 97.5) will closer to the theoretical value.


# Question 6

## a)

```{r}
data <- read_table('vote.txt')
ggplot(data,aes(x=growth, y=vote)) + geom_point() + geom_smooth(formula = y ~ x, method=lm, se=FALSE)
```

## b)

```{r}
lm <- lm(vote ~ growth , data)
summary(lm)
```
$\hat{\beta_0}$ = 46.2476.It means if the average personal income growth in the pervious year is 0, the incumbent party's share of the two-party vote will be 46.2476 percent.

$\hat{\beta_1}$ = 3.0605.It means as the average personal income growth in the pervious year increasing one unit, the incumbent party's share of the two-party vote will increase 3.0605 percent.


## c)

In R:
```{r}

predict(lm, data.frame(growth = 0.1))
```

By hands: 
```{r}
data
```
In 2018, the growth x is 0.1. 

$\hat{Y}$ = $\hat{\beta_0}$ + $\hat{\beta_1}*X$ = 46.2476 + 3.0605 * 0.1 = 46.5536

"Obama" would win. Because the vote rate for Obama is 54.4464 and the vote rate for Mccain is "46.5536".46.5536 is smaller than 54.4464(1- 46.5536).



## d)

$H_0 :\beta_1 = 0$

$H_1 :\beta_1 > 0$
  

```{r}
summary(lm)
1-pt(4.396,14)
```
Therefore, the test statistics is 4.396.

n = 16, n-2 = 14.

p-value: $P(t_{14} > 4.396)$ = 0.0003047443

p-value < 0.05, p-value < $\alpha$. Reject $H_0$.

There is a positive association between incumbent party's vote share and economical growth.


## e)

By hands :
```{r}
qt(0.975, 14)
```
95% CI for $\beta_1$ : ( $\hat{\beta_1}$ - $t_{0.975,n-2} * se(\hat{\beta_1})$ , $\hat{\beta_1}$ + $t_{0.975,n-2} * se(\hat{\beta_1})$ )

$\hat{\beta_1}$ = 3.0605

$t_{0.975,n-2}$ = 2.144787

$se(\hat{\beta_1})$ = 0.6963

3.0605 - 2.144787 * 0.6963 = 1.567085

3.0605 - 2.144787 * 0.6963 = 4.553915

Answer:(1.567085, 4.553915)

By R:
```{r}
confint(lm, level=0.95)
```
Answer:(1.567169, 4.553887)


## f)

## Step 1:depart the absolute sign

$p(|\hat{\beta_1} - \beta| > 1)$ = $p(\hat{\beta_1} - \beta > 1)$ + $p(\hat{\beta_1} - \beta < -1)$

## Step 2:multiply both side by 
                                $se(\hat{\beta_1}):\ $
                                
$P(\frac{\hat{\beta_1} - \beta}{se(\hat{\beta_1})}>\frac{1}{se(\hat{\beta_1})})$ + $P((\frac{\hat{\beta_1} - \beta}{se(\hat{\beta_1})}<-\frac{1}{se(\hat{\beta_1})})$

## Recall the test-statistic formula and using the porperty to find a more easy-handle data:
$P(\frac{\hat{\beta_1} - \beta}{se(\hat{\beta_1})}>\frac{1}{se(\hat{\beta_1})}) = P(t^*(different\ than \ the\ t^*\ showed\ in (b)))>\frac{1}{se(\hat{\beta_1})})$
$P((\frac{\hat{\beta_1} - \beta}{se(\hat{\beta_1})}<-\frac{1}{se(\hat{\beta_1})})= P(t^*(different\ than \ the\ t^*\ showed\ in (b)))<-\frac{1}{se(\hat{\beta_1})})$

## with this data, we can compare to the p-value and using diagram to see the region:
$since\ \frac{1}{se(\hat{\beta_1})}\ >\ 0\ ->\ P(t^*>\frac{1}{se(\hat{\beta_1})})\ is\ showing\ below\ :$
```{r}
sD <- 2;
x <- seq(-5,5,by=0.1)
y <- dnorm(x,sd=sD)
right <- 2.5631031310892
tail <- qnorm(0.9,sd=sD)
plot(x,y,type="l",xaxt="n",ylab="",xlab=expression(paste('')),
     axes=FALSE,ylim=c(0,max(y)*1.05),xlim=c(min(x),max(x)),
       frame.plot=FALSE)
axis(1,at=c(-5,right,0,5),labels=c(expression(' '),expression(1/SE(beta[1])),expression(' '),expression(' ')))
xtail <- seq(right,6,by=0.1)
ytail <- dnorm(xtail,sd=sD)
polygon(c(xtail,xtail[length(xtail)],xtail[1]),
          c(ytail, 0, 0), col='red')
```

$since\ \frac{1}{se(\hat{\beta_1})}\ >\ 0\ ->\ P(t^*>-\frac{1}{se(\hat{\beta_1})})\ is\ showing\ below\ :$
```{r}
sD <- 2;
x <- seq(-5,5,by=0.1)
y <- dnorm(x,sd=sD)
tail <- qnorm(0.9,sd=sD)
plot(x,y,type="l",xaxt="n",ylab="",xlab=expression(paste('')),
     axes=FALSE,ylim=c(0,max(y)*1.05),xlim=c(min(x),max(x)),
       frame.plot=FALSE)
axis(1,at=c(-5,-right,0,5),labels=c(expression(' '),expression(-1/SE(beta[1])),expression(' '),expression(' ')))
xtail <- seq(right,5,by=0.1)
ytail <- dnorm(xtail,sd=sD)
polygon(c(-xtail,-xtail[length(xtail)],-xtail[1]),
          c(ytail, 0, 0), col='red')
```

## Based on the comparison and the two diagrams, we can conclude it is just :
```{r}
pt(1/0.6963,14,lower.tail = FALSE)*2
```

Answer: 0.1729238












